---
title: "`select()` and `pull()`"
subtitle: ""
date: ""
author: "Dr. Kao"
format:
  html:
    embed-resources: true
    code-overflow: wrap
    number-sections: true
---

## Load Packages and Data

```{r}
#| message: false
library()

library()

library()

library()

dogs <- read_csv() |>
  clean_names()

earthquakes <- read_csv() |>
  clean_names()

fastfood <- read_csv() |>
  clean_names()

water_quality <- read_csv() |> 
  clean_names()
```

## Introduction 

There are two verbs in the `dplyr` grammar that let us extract column(s) from a data frame. 
* `pull()` extracts a single column from a data frame and returns it as a vector, similar to the `$` operator in base R.
* `select()` extracts multiple columns from a data frame and returns the result as a data frame.
* Both function can take column names or column positions as arguments.

## `pull()`

:::{.callout-note icon='false'}
## Example 1.

Extract the `water_temperature_c` column from the `water_quality` data frame.

_Solution._

First, use the `names()` function to check the column names of the `water_quality` data frame:

```{r}

```

**Method 1:** `pull` using the column name.

```{r}
water_temp_c_1 <- 
```

```{r}
head()

class()
```

**Method 2:** `pull` using the column position. The output from `names(water_quality)` shows us that the `water_temperature_c` is the 7th column in the data frame.

```{r}
water_temp_c_2 <- 
```

```{r}
head()

class()
```
:::

:::{.callout-note icon='false'}
## Example 2.

Extract the last column from the `water_quality` data frame. 

_Solution._

We can use the `names()` to see that the name of the last column is `longitude`, then use `pull()` to extract that column by name.

Alternatively, we can use `-1` as the column position to achieve the same result.

```{r}
longitude <- 
```

```{r}
head()
```
:::

:::{.callout-warning icon='false'}
## Remark.

`pull()` and `$` do almost exactly the same thing, but `pull()` is preferred in our case because conforms to our code involving pipes.
:::

## `select()`

:::{.callout-note icon='false'}
## Example 2.

Extract the `flipper_length_mm` and `body_mass_g` columns from the `penguins` data frame.

_Solution._

**Method 1.** Using the column names.

```{r}
penguins_fl_bm_1 <- 
```

```{r}
head()

class()
```

**Method 2.** Using the column positions.

```{r}
penguins_fl_bm_2 <- 
```

```{r}
head()

class()
```
:::

## Appendix

Belwo is a list of new functions introduced in this section.

| Function   | Description                                         |
|------------|-----------------------------------------------------|
| `pull()`   | Extract a single column as a vector, similar to `$` |
| `select()` | Extract columns as data frame                       |
: {.striped}