---
title: "Data Frames in R"
subtitle: "CIS 351: Week X Day Y"
date: ""
author: ""
format:
  html:
    embed-resources: true
    code-overflow: wrap
    number-sections: true
---

## Load Packages and Data

```{r}
#| message: false
library()

library()

library()
```

## Data Types in R

There are three main types of data we will work with in this course

* ****: The usual numbers
* ****: Letters; these can be mixed numbers or symbols
* ****: Binary values corresponding to `TRUE` or `FALSE`

:::{.callout-important icon='false'}
# The `class()` Function

We can determine the data type of an object by using the `class()` function. 
:::

### Numeric Values

```{r}
class()
```

```{r}
class()
```

```{r}
bob <- 

class()
```

### Character Values

Character values are typically values that are wrapped around by `''` or `""`.

```{r}

```

```{r}
class()
```

```{r}

```

```{r}
class()
```

```{r}
my_name <- 
```

```{r}
class(my_name)
```

### Logical Values

There are only two logical values: `TRUE` and `FALSE`. 

```{r}

```

```{r}

```

```{r}
class()

class()
```

```{r}

```

```{r}

```

## `NA` Values

`NA` (not available) is a special type of data in R. 

```{r}

```

```{r}
class()
```
```{r}

```

```{r}

```