---
title: "The `mutate()` Function"
subtitle: ""
date: ""
author: ""
format:
  html:
    embed-resources: true
    code-overflow: wrap
    number-sections: true
---

## Load Packages and Data

```{r}
#| message: false
library(tidyverse)

library(janitor)

library(readxl)

library(palmerpenguins)

dogs <- read_csv() |>
  clean_names()

earthquakes <- read_csv() |>
  clean_names()

fastfood <- read_csv() |>
  clean_names()

water_quality <- read_csv() |> 
  clean_names()
```

## Adding and Modifying Columns 

### The `mutate()` function

The `mutate()` function can add and modify existing columns in a data frame. 

:::{.callout-note icon='false'}
## Example 1.

Use the `mutate()` function to add a new column to the `dogs` data frame that contains the dogs' weights in kilograms (kg). Round all final values to one decimal place.

_Solution._ First recall that 1 lb is approximately equal to 0.45359237 kg.

```{r}

```
:::

:::{.callout-note icon='false'}
## Example 2.

Convert the water temperature to °F in the `water_quality` data frame. Drop the original temperature column and move the new temperature column to its position.

Recall that the formula to convert from °C to °F is
$$
  F = \frac{9}{5} C + 32.
$$

_Solution._ The optional `.keep` argument of `mutate()` lets use control whether the columns used in `mutate()` are retained or removed. The `.before` and `.after` arguments allow us to specify where the new column or columns should be places.

```{r}

```

```{r}

```
:::

### Group-wise calculations with `mutate()`

Many `dplyr` functions have the powerful capability of performing calculations within groups. This is typically achieved by specifying the `.by` argument and providing one or more variables.

:::{.callout-note icon='false'}
## Example 3.

Add a new column to the `dogs` data where the new column shows the mean weight (in pounds) within each breed. 

```{r}

```
:::

:::{.callout-note icon='false'}
## Example 4.

Continuing from Example 4, add a new column to the `dogs_1` data frame that compares each dog's weight to the mean weight of its breed using the following rules:

* If the dog's weight is greater than the mean breed weight, assign "heavy" to the new column.
* If the dog's weight is less than the mean breed weight, assign "light" to the new column.
* If the dog's weight is equal to the mean breed weight, assign "average" to the new column.

_Solution._ Whenever we need to create new values based on one or more conditions, `case_when()` is an extremely useful helper function. It provides a cleaner and more readable alternative to traditional `if...else` statements in most programming languages.

**Method 1:**
```{r}

```

**Method 2:**
```{r}

```
:::

## Appendix

Belwo is a list of new functions introduced in this section.

| Function       | Description                                         |
|----------------|-----------------------------------------------------|
| `mutate()`     | Creates new columns or modify existing columns      | 
| `case_when()`  | Assigns values based on multiple conditions         | 
: {.striped}