---
title: "Summary Statistics for Vectors in R"
subtitle: "CIS 351:"
date: today
author: ""
format:
  html:
    embed-resources: true
    code-overflow: wrap
    number-sections: true
---

## Load Packages and Data

```{r}
#| message: false
library(tidyverse)

library(janitor)

library(readxl)

library(palmerpenguins)

dogs <- read_csv("/Users/chriskao/Documents/CIS351/Week 3/dogs.csv") |>
  clean_names()

earthquakes <- read_csv("/Users/chriskao/Documents/CIS351/Week 3/earthquakes.csv") |>
  clean_names()

fastfood <- read_csv("/Users/chriskao/Documents/CIS351/Week 3/fastfood.csv") |>
  clean_names()
```

## Summary Statistics of Vectors in R
### Summary Functions

Base R and the `{dplyr}` package have various functions that can provide summaries of a vector. Below is a table of frequently used summary functions in R.

| Function            | R Code     | Description                             |
|---------------------|------------|-----------------------------------------|
| Length              | `length()` | Counts the number of values in a vector |
| Distinct Values     | `unique()` | Finds all distinct values in a vector     |
| Number of Distinct Values | `n_distinct()` | Count the number of distinct values in a vector
| Sum                 | `sum()`    | Adds all numeric values together        |
| Mean                | `mean()`   | Calculates the average                  |
| Median              | `median()` | Finds the median value                  |
| Standard Deviation  | `sd()`     | Calculates the standard deviation       |
| Variance            | `var()`    | Calculate the variance                  |
| Minimum             | `min()`    | Finds the minimum value                 |
| Maximum             | `max()`    | Finds the maximum value                 |
| Interquartile Range | `IQR()`    | Calculates the interquartile range      |
: {.striped}

:::{.callout-note icon='false'}
## Example 1.

a. Find the number of observations in the `name` variable of the `dogs` data frame.
b. Find the number of observations in the `island` variable of the `penguins` data frame.

_Solutions._

```{r}
# (a)
length(dogs$name)
```

```{r}
# (b)
length(penguins$island)
```
:::

:::{.callout-note icon='false'}
## Example 2.

a. Find all unique dog names and breeds in the `dogs` data frame.
b. Find all unique penguin species and sex in the `penguins` data frame.

_Solutions._

```{r}
# (a)
unique(dogs$name)

unique(dogs$breed)
```

```{r}
# (b)
unique(penguins$species)

unique(penguins$sex)
```
:::

:::{.callout-note icon="false"}
## Example 3.

a. How many unique dog names and breeds are there in the `dogs` data frame?
b. How many unique penguin species and sex are there in the `penguins` data frame?

_Solutions._

```{r}
# (a)
n_distinct(dogs$name)

n_distinct(dogs$breed)
```

```{r}
# (b)
n_distinct(penguins$species)

n_distinct(penguins$sex) # this counts the NA

n_distinct(penguins$sex, na.rm = TRUE)
```
:::

:::{.callout-note icon='false'}
## Example 4.

a. Find the total weight of all dogs in the `dogs` data frame.
b. Find the total flipper length of all penguins in the `penguins` data frame.

_Solutions._

```{r}
# (a)
sum(dogs$weight_lbs)
```

```{r}
# (b)
sum(penguins$flipper_length_mm)

sum(penguins$flipper_length_mm, na.rm = TRUE)
```
:::

:::{.callout-note icon='false'}
## Example 5.

Find the following summary statistics for the weights of all dogs in the `dogs` data frame:
  a. Mean
  b. Median
  c. Standard deviation
  d. Variance
  e. Minimum
  f. Maximum
  g. Interquartile range

_Solutions._

```{r}
# (a)
mean(dogs$weight_lbs)
```

```{r}
# (b)
median(dogs$weight_lbs)
```

```{r}
# (c)
sd(dogs$weight_lbs)
```

```{r}
# (d)
var(dogs$weight_lbs)
```

```{r}
# (e)
min(dogs$weight_lbs)
```

```{r}
# (f)
max(dogs$weight_lbs)
```

```{r}
# (g)
IQR(dogs$weight_lbs)
```
:::

:::{.callout-note icon='false'}
## Example 6.

Repeat Example 5 using the penguins' body mass.

_Solutions._

```{r}
# (a)
mean(penguins$body_mass_g, na.rm = TRUE)
```

```{r}
# (b)
median(penguins$body_mass_g, na.rm = TRUE)
```

```{r}
# (c)
sd(penguins$body_mass_g, na.rm = TRUE)
```

```{r}
# (d)
var(penguins$body_mass_g, na.rm = TRUE)
```

```{r}
# (e)
min(penguins$body_mass_g, na.rm = TRUE)
```

```{r}
# (f)
max(penguins$body_mass_g, na.rm = TRUE)
```

```{r}
# (g)
IQR(penguins$body_mass_g, na.rm = TRUE)
```
:::

### Calculating Quantiles

:::{.callout-warning icon='false'}
## Quartiles and Quantiles

A **quantile** is a value that identifies a specified percentage of observations in an ordered data set. **Quartiles** and **percentiles** are special cases of quantiles that divide the data into four and 100 equal parts, respectively.

In R, the `quantile()` returns the five-number summary by default. You can also specify percentage values to find quantiles at other locations in the data
:::

:::{.callout-note icon='false'}
## Example 7.

Find the five-number summary for the penguins' flipper lengths.

_Solution._
```{r}
quantile(penguins$flipper_length_mm, na.rm = TRUE)
```
:::

:::{.callout-note icon='false'}
## Example 8.

Calculate the 10th, 30th, 50th, 70th, and 90th percentiles of the penguins’ bill lengths.

_Solution._
```{r}
quantile(
  penguins$bill_length_mm, # specify the variable
  probs = c(0.1, 0.3, 0.5, 0.7, 0.9), # specify the percentiles
  na.rm = TRUE # remove NAs from the calculation
)
```
:::

### The `summary()` Function

The `summary()` function is a base R function that calculates the five-number summary, along with the mean and the number of NA's in a numeric variable. When applied to a character variable, the `summary()` function returns the number of observations, the number of of distinct values, and the number of NA's.

:::{.callout-caution}
The `summary()` function does not provide the standard deviation, variance, or interquartile range of a numeric variable.
:::

:::{.callout-note icon='false'}
## Example 9.

Use the `summary()` function to find the five-number summary of penguins' bill depths.

_Solution._

```{r}
summary(penguins$bill_depth_mm)
```
:::

:::{.callout-note icon='false'}
## Example 10.

a. Use the `summary()` function on the `breed` variable in the `dogs` data frame.
b. Use the `summary()` function on the `region` variable in the `earthquakes` data frame.

_Solutions._

```{r}
# (a)
summary(dogs$breed)
```

```{r}
# (b)
summary(earthquakes$region)
```
:::
