---
title: "Indexing and Subsetting in R"
subtitle: "CIS 351:"
date: today
author: ""
format:
  html:
    embed-resources: true
    code-overflow: wrap
    number-sections: true
---

## Load Packages and Data

```{r}
#| message: false
library(tidyverse)

library(janitor)

library(readxl)

library(palmerpenguins)

dogs <- read_csv("/Users/ckao/Library/CloudStorage/OneDrive-FlaglerCollege/CIS351/2026 Summer A/Notes/dogs_2.csv") |>
  clean_names()

earthquakes <- read_csv("/Users/ckao/Library/CloudStorage/OneDrive-FlaglerCollege/CIS351/2026 Summer A/Notes/earthquakes.csv") |>
  clean_names()

fastfood <- read_csv("fastfood.csv") |>
  clean_names()
```

## The Index Operators in R

In most modern programming languages, indexing operators are built in, allowing us to access elements in arrays, vectors, and similar data structures. 

### The `$` operator

We have already seen that the `$` operator can be used to extract columns from a data frame in R.

:::{.callout-note icon='false'}
## Example 1.

Extract the `restaurant` column from the `fastfood` data frame. What are the unique items in this column? How many unique items are there?

_Solution._

```{r}
restaurant <- fastfood$restaurant
```

```{r}
unique(restaurant)
```

```{r}
n_distinct(restaurant)
```
:::

### The `[]` operator

The `[]` operator in R is used for **subsetting** objects by extracting or selecting elements from vectors, matrcies, data frames, and other objects. It returns a subset of the original object based on specified row and/or column indices, names, or logical conditions.

:::{.callout-note icon='false'}
## Example 1. (Subsetting by Indices)

Consider the vector

```{r}
name <- dogs$name
```

a. What is the 7th element in the `name` vector?
b. What are 3rd through the 11th elements in the `name` vector?
c. What are the 4th through the last element in the `name` vector?
d. Remove the 1st element from the `name` vector.
e. Remove the 1st and the 12th elements from the `name` vector.
_Solutions._

```{r}
# (a)
name[7]
```

```{r}
# (b)
name[3:11]
```

```{r}
# (c)
name[4:length(name)]
```

```{r}
# (d)
name[-1]
```

```{r}
# (e)
name[-c(1, 12)]
```
:::
