---
title: "Logical Subsetting in R"
subtitle: ""
date: ""
author: ""
format:
  html:
    embed-resources: true
    code-overflow: wrap
    number-sections: true
---

## Load Packages and Data

```{r}
#| message: false
library()

library()

library()

library()

dogs <- read_csv() |>
  clean_names()

earthquakes <- read_csv() |>
  clean_names()

fastfood <- read_csv() |>
  clean_names()
```

## Logical Subsetting in R

Previously, we discussed how to use relational and logical operations to determine elements in a vector that satisfy a certain condition or a set of conditions. 

:::{.callout-note icon='false'}
## Example 0.

Consider the following vector of dog weights from the `dogs` data frame. Which weights are greater than the mean weight?

```{r}
weight_lbs <- 
```

_Solution._

```{r}
weight_lbs > 

heavy_indices <- 
```
:::

In Example 0, we saw that the relational operation only returns a logical vector indicating which elements satisfy the condition. Rather than viewing only these `TRUE` and `FALSE` values, we often want to use the logical vectors to retain only the elements that satisfy a condition or a set of conditions. 

:::{.callout-note icon='false'}
## Example 2. 

Determine the values in `weight_lbs` that are greater than the mean dog weight.

_Solution._ We will present three solutions, each representing a higher level of abstraction than the previous one.

**Method 1.** The output of the code `weight_lbs > mean(weight_lbs, na.rm = TRUE)` is a vector of logical values. Let us pass this logical vector to the `[]` operator for `weight_lbs`.

```{r}
c(TRUE, TRUE, TRUE, FALSE, TRUE, TRUE, FALSE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE)
```

**Method 2.** Since we defined the logical vector as `heavy_indices`, we can pass this object to the `[]` operator instead of typing out the entire vector.

```{r}
weight_lbs[]
```

**Method 3.** We can dispense with assigning the result of Example 0 to a variable and pass the relational statement to the `[]` operator to obtain the same result.

```{r}
weight_lbs[]
```
:::

:::{.callout-note icon='false'}
## Example 3.

From the `earthquakes` data frame, determine the earthquake magnitudes that fall between the first quartile (exclusive) and the third quartile (inclusive).

_Solution._

```{r}
richter <- 

richter[]
```
:::