---
title: "Data Frames in R"
subtitle: ""
date: 
author: ""
format:
  html:
    embed-resources: true
    code-overflow: wrap
    number-sections: true
---

## Load Packages and Data

```{r}
#| message: false
library()

library()

library()

library()

dogs <- read_csv() |>
  clean_names()

earthquakes <- read_csv() |>
  clean_names()

fastfood <- read_csv() |>
  clean_names()
```

## Data Frames in R

We have encountered data frames many times throughout this course. A **data frame** is a collection of vectors of equal length, where each vector represents a variable. 

:::{.callout-note icon='false'}
## Example 1.

Construct a data frame in R using the following table.

| Name            | Breed              | Weight (lbs) |
|-----------------|--------------------|--------------|
| Bella           | Labrador Retriever |           65 |
| Max             | German Shepard     |           75 |
| Carter          | Mixed              |           68 |
| Daisy           | Beagle             |           22 |
: {.striped}

_Solution._

To construct a data frame in R, we use the `data.frame()` function and provide the column vectors as its arugments.

```{r}
dogs_abbr <- data.frame()

dogs_abbr
```
:::

The following functions can be used to obtain basic information about a data frame in R.

| Function     | R Code      | Description                                        |
|--------------|-------------|----------------------------------------------------|
| Row Count    | `nrow()`    | Returns the number of rows                         |
| Column Count | `ncol()`    | Returns the number of columns                      |
| Dimensions   | `dim()`     | Returns the dimensions of the data frame           |
| Column Names | `names()`   | Returns a vector of column names                   |
| Structure    | `str()`     | Displays an internal structure of the data frame   |
| Summary      | `summary()` | Returns a basic statistical summary of all columns |
: {.striped}

:::{.callout-note icon='false'}
## Example 2.

Apply each of the functions from the table above to the `fastfood` data frame.

_Solution._

```{r}
nrow()
```

```{r}
ncol()
```

```{r}
dim()
```

```{r}
names()
```

```{r}
str()
```

```{r}
summary()
```
:::