---
title: "Vectorized Operations in R"
subtitle: ""
date: ""
author: ""
format: 
  html:
    embed-resources: true
    code-overflow: wrap
---

## Load Packages and Data

```{r}
#| message: false
 
library()

library()

library()

library()
```

## Relational Operations

The **relational operations** in R compare values and return logical results: `TRUE` or `FALSE`. The symbols used to compare the values are called the **relational operators**. These operators are summarized in the table below.

```{css, echo=FALSE}
.center-table table {
  width: var(--table-width, 60%);
  margin-left: auto;
  margin-right: auto;
}
```

:::{.center-table}
| Operator Name            | Mathematical Symbol | R Code |
|--------------------------|:-------------------:|:------:|
| Equal to                 | $=$                 | `==`   |
| Not equal to             | $\ne$               | `!=`   |
| Less than                | $<$                 | `<`    |
| Less than or equal to    | $\leqslant$         | `<=`   |
| Greater than             | $>$                 | `>`    |
| Greater than or equal to | $\geqslant$         | `>=`   |
: {.striped .hover}
:::

:::{.callout-note icon='false'}
## Example 1.

Answer the following questions using the `weight_lb` vector defined below. 

```{r}
weight_lb <- c(65, 75, 68, 22, 40, 50, 20, 60, 35, 22)
```

a. Which values in the vector are equal to 40? 
b. Which values are equal to 77?
c. Which values are _not_ equal to 22?
d. Which values are less than 65?
e. Which values are less than or equal to 70?
f. Which values are greater than 50?
g. Which values are greater than or equal to 50?

_Solutions._
```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```
:::

## Logical Operations

**Logical operations** in R are used to combine or negate logical conditions that evaluate to `TRUE` or `FALSE`. These operations are especially useful when used with relational operators to filter data, evaluate multiple conditions, and control program flow. In the expressions below, `X` and `Y` represent logical statements or conditions that return either `TRUE` or `FALSE`. The logical operators available in R are summarized in the table below.

:::{.center-table style="--table-width: 75%;"}
| Operator Name | R Code |   R Syntax  | Description                                         |
|---------------|:------:|:-----------:|-----------------------------------------------------|
| And           |   `&`  |   `X & Y`   | Both `X` and `Y` must be true                       |
| Or            |  `\|`  |   `X \| Y`  | Either `X` is true or `Y` is true, or both are true |
| Exclusive or  |  `xor` | `xor(X, Y)` | Either `X` is true or `Y` is true, but not both     |
| Negation      |   `!`  |     `!X`    | Negates the truth value of `X`                      |
: {.striped .hover}
:::

:::{.callout-note icon='false'}
## Example 2.

Use the `weight_lb` vector from Example 1 to complete the following questions.

a. Determine the values that are greater than 50 and less than or equal to 70.
b. Determine the values that are greater than or equal to 50 or less than 70.
c. Determine the values that are great than 50 xor less than 70.
d. Negate the truth value of parts a, b, and, c.

_Solutions._

```{r}

```

```{r}

```

```{r}

```

```{r}

```
:::

## The `%in%` Operator

The `%in%` operator in R is used to determine whether elements of one vector are contained in another vector. It returns a logical vector of `TRUE` or `FALSE` values indicating whether each element of the vector on the left side of `%in%` belongs in the vector on the right side of `%in%`.

:::{.callout-note icon='false'}
## Example 3.

Use the `%in%` operator to determine whether the values 20 and 33 are in `weight_lb`.

_Solutions._

**Method 1.**
```{r}

```

**Method 2.**
```{r}

```
:::